Zero-shot task adaptation by homoiconic meta-mapping

### 2019-11-09 writing/homoiconic_meta_mapping_appendix.tex
Further tweaks from Jay
More reviewer suggestions

@@

\item On each basic task, some of the data ($D_1$) is fed to the meta-network $\mathcal{M}$ while some ($D_2$) is held out. This encourages the model to actually infer the underlying function, rather than just memorizing the examples.

@@

\subsection{Source repositories}
The full code for the experiments and analyses can be found on github:
%%\begin{itemize}
%%\item \url{https://github.com/lampinen/polynomials}
%%    \begin{itemize}
%%        \item See the \verb|continual_learning| and \verb|language_meta_only| branches for the continual learning and language results, respectively. 
%%    \end{itemize}
%%\item \url{https://github.com/lampinen/meta_cards} 
%%\end{itemize}
(will be made available in the de-anonymized version.)
\section{Numerical results} \label{app_numerical_results}
In this section we provide the mean values and bootstrap confidence intervals corresponding to the major figures in the paper, as well as the baseline results in those figures. Tables were generated with stargazer \citep{Hlavac2018}. \par

>>>>

\item On each basic task, some of the data ($D_1$) is fed to the meta-network as examples $\mathcal{M}$ while some probes ($D_2$) is held out. This encourages the model to actually infer the underlying function, rather than just memorizing the examples. However, the probe labels are still used in training the system end-to-end, in order to encourage it to generalize.

@@

@@

\subsection{Source repositories}
The full code for the experiments and analyses can be found on github:
\begin{itemize}
\item \url{https://github.com/lampinen/polynomials}
    \begin{itemize}
        \item See the \verb|continual_learning| and \verb|language_meta_only| branches for the continual learning and language results, respectively. 
    \end{itemize}
\item \url{https://github.com/lampinen/meta_cards} 
\end{itemize}
\section{Numerical results} \label{app_numerical_results}
In this section we provide the mean values and bootstrap confidence intervals corresponding to the major figures in the paper, as well as the baseline results in those figures. Tables were generated with stargazer \citep{Hlavac2018}. \par

### 2019-11-09 writing/homoiconic_meta_mapping_main.tex
Tweaks
More reviewer suggestions

\vspace{-0.25em}
\section{Homoiconic meta-mapping (HoMM) architecture}
\vspace{-0.5em} % section
To address these challenges, we propose HoMM architectures, composed of two components: 
\vspace{-0.5em}
\begin{enumerate} \setlength \itemsep{0em}
\item Input/output systems: domain specific encoders and decoders (vision, language, etc.) that map into a shared embedding space $Z$.
\item A meta-learning system that a) learns to embed tasks into the shared embedding space $Z$, b) learns to use these task embeddings to perform task-appropriate behavior, c) learns to embed meta-mappings into the same space, and d) learns to use these meta-mapping embeddings to transform basic task embeddings in a meta-mapping appropriate way.\end{enumerate}
\vspace{-0.5em}
\looseness=-1
These architectures are homoiconic because they have a completely shared $Z$ for individual data points, tasks, and meta-mappings. Why is this useful? The primary advantage is that it parsimoniously allows for arbitrary mappings between these entities. In addition to basic tasks, the system can learn to perform meta-mappings to follow instructions or change behavior. That is, it can transform task representations using the same components it uses to transform basic data points. (See also appendix \ref{app_lesion_results_shared_z}.) \par
Without training on meta-mappings, of course, the system will not be able to execute them well. However, as we will show, if it is trained on a broad enough set of such mappings, it will be able to generalize to new instances drawn from the same meta-mapping distribution. For instances that fall outside its data distribution, or for optimal performance, it may require some retraining, however. This reflects the structure of human behavior -- we are able to adapt rapidly when new knowledge is relatively consistent with our prior knowledge, but learning an entirely new paradigm (such as calculus for a new student) can be quite slow \citep[cf.][]{Kumaran2016, Botvinick2019}. \par 
\begin{figure}[t]
\centering
\begin{tikzpicture}[scale=0.8, every node/.style={scale=0.8}]
%% basic meta learning
\begin{scope}[shift={(0.4, 0.4)}]
\draw[boundingbox, fill=white] (-3, -4.3) rectangle (2.5, 3.2);
\end{scope}
\begin{scope}[shift={(0.2, 0.2)}]
\draw[boundingbox, fill=white] (-3, -4.3) rectangle (2.5, 3.2);
\end{scope}

@@

\paragraph{Basic meta-learning:} To perform a basic task, input and target encoders ($\mathcal{I}$ and $\mathcal{T}$) are used to embed individual pairs from an example dataset \(D_1\), to form a dataset of example (input, output) tuples (fig. \ref{architecture_inference_fig}a). These examples are fed to $\mathcal{M}$, which produces a function embedding (via a deep neural network, with several layers of parallel processing across examples, followed by an element-wise max across examples, and several more layers). This function embedding is mapped through the hyper network $\mathcal{H}$ to parameterize $F$, and then $F$ is used to process a dataset of embedded probe inputs, and $\mathcal{O}$ to map the resultant embeddings to outputs. This system can be trained end-to-end on target outputs for the probes. Having two distinct datasets forces generalization at the meta-learning level, see appendix \ref{app_clarifying_holdouts}. See appendix \ref{app_model_details} for detailed architecture, and hyper-parameters. \par 

@@

The system can then be trained end-to-end on this loss to adjust the weights of \(\mathcal{T,H,M,O}\), and \(\mathcal{I}\).

@@

\paragraph{Meta-mapping:} The fundamental insight of our paper is to show how basic tasks and meta-mappings can be treated homogenously, by allowing the network to transform its task representations like data (see fig. \ref{architecture_inference_fig}b,c). From the perspective of our architecture, learning a meta-mapping between tasks is exactly analogous to learning a basic task. Anything that is embedded in $Z$ can be transformed using the same system. Because tasks are embedded in $Z$ for basic meta-learning, this allows for meta-mappings using exactly the same $\mathcal{M}$ and $\mathcal{H}$ that we use for basic tasks. Just as we would take a set of paired embeddings of data points for a basic task, and use them to compute a function embedding for that task, we can take a set of paired function embeddings, and use them to create an embedding for the meta-mapping. We can then use this meta-mapping embedding to transform another task. We can thus behave zero-shot on a novel task based on its relationship to a prior task. \par
For example, suppose we have an embedding $z_{game1} \in Z$ for the task of playing some game, and we want to switch to trying to lose this game. We can generate a meta-mapping embedding $z_{\text{meta}} \in Z$ from examples of embeddings generated by the system when it is trying to win and lose various games: $z_{\text{meta}} = \mathcal{M}\left( \left\{\left((z_{game2},z_{game2,lose}\right), ... \right\}\right)$. We can generate a new task embedding $\hat{z}_{game1,lose} \in Z$:  

@@

\paragraph{Related work:} There has been a variety of past work on zero-shot learning, mostly based on natural-language descriptions of tasks. For example, \citet{Larochelle2008} considered the general problem of behaving in accordance with language instructions as simply asking a model to adapt its response when conditioned on different ``instruction'' inputs. Later work explored zero-shot classification based on only a natural language description of the target class \citep{Socher2013,Romera-Paredes2015,Xian2018}, or of a novel task in a language-conditioned RL system \citep{Hermann2017}. Some work has even exploited relationships between tasks as a learning signal \citep{Oh2017a, Pal2019}. However, to our knowledge none of the prior work has proposed using meta-mapping approaches to adapt to new tasks. \par

>>>>

\vspace{-0.25em}
\section{Homoiconic meta-mapping (HoMM) architecture}
\vspace{-0.5em} % section
To address these challenges, we propose HoMM architectures, composed of two components: 
\vspace{-0.5em}
\begin{enumerate} \setlength \itemsep{0em}
\item Input/output systems: domain specific encoders and decoders (vision, language, etc.) that map into a shared embedding space $Z$.
\item A meta-learning system that a) learns to embed tasks into the shared embedding space $Z$, b) learns to use these task embeddings to perform task-appropriate behavior, c) learns to embed meta-mappings into the same space, and d) learns to use these meta-mapping embeddings to transform basic task embeddings in a meta-mapping appropriate way.\end{enumerate}
\vspace{-0.5em}
\looseness=-1
These architectures are homoiconic because they have a completely shared $Z$ for individual data points, tasks, and meta-mappings. This parsimoniously allows for arbitrary mappings between these entities. In addition to basic tasks, the system can learn to perform meta-mappings to systematically alter task behavior based on relationships to other tasks. That is, it can transform task representations using the same components it uses to transform basic data points. (See also appendix \ref{app_lesion_results_shared_z}.) \par
Without training on meta-mappings, of course, the system will not be able to execute them well. However, as we will show, if it is trained on a broad enough set of such mappings, it will be able to generalize to new instances drawn from the same meta-mapping distribution. For instances that fall outside its data distribution, or for optimal performance, it may require some retraining, however. This reflects the structure of human behavior -- we are able to adapt rapidly when new knowledge is relatively consistent with our prior knowledge, but learning an entirely new paradigm (such as calculus for a new student) can be quite slow \citep[cf.][]{Kumaran2016, Botvinick2019}. \par 
\begin{figure}[t]
\centering
\begin{tikzpicture}[scale=0.8, every node/.style={scale=0.8}]
%% basic meta learning
\begin{scope}[shift={(0.4, 0.4)}]
\draw[boundingbox, fill=white] (-3, -4.3) rectangle (2.5, 3.2);
\end{scope}
\begin{scope}[shift={(0.2, 0.2)}]
\draw[boundingbox, fill=white] (-3, -4.3) rectangle (2.5, 3.2);
\end{scope}

@@

\paragraph{Basic meta-learning:} To perform a basic task, input and target encoders ($\mathcal{I}$ and $\mathcal{T}$) are used to embed individual pairs from an example dataset \(D_1\), to form a dataset of example (input, output) tuples (fig. \ref{architecture_inference_fig}a). These examples are fed to $\mathcal{M}$, which produces a function embedding (via a deep neural network, with several layers of parallel processing across examples, followed by an element-wise max across examples, and several more layers). This function embedding is mapped through the hyper network $\mathcal{H}$ to parameterize $F$, and then $F$ is used to process a dataset of embedded probe inputs, and $\mathcal{O}$ to map the resultant embeddings to outputs. This system can be trained end-to-end on target outputs for the probes. Having two distinct datasets forces generalization at the meta-learning level, in exactly the same way as standard meta-learning algorithms learn to generalize. \par 

@@

The system can then be trained end-to-end on this loss to adjust the weights of \(\mathcal{T,H,M,O}\), and \(\mathcal{I}\). That is, the meta-learning training procedure optimizes the system for generalizing the function implied by the examples it sees, in order to respond appropriately to probe inputs. As in training a standard meta-learning model, these probes are held-out from the meta-network, but are still provided during training. Other examples are held-out entirely during training, and only used for evaluation, see appendix \ref{app_clarifying_holdouts} for some detailed description. See appendix \ref{app_model_details} for details on the architecture and hyper-parameters. \par

@@

\paragraph{Meta-mapping:} The fundamental insight of our paper is to show how basic tasks and meta-mappings can be treated homogenously, by allowing the network to transform its task representations like data (see fig. \ref{architecture_inference_fig}b,c). That is, we propose to transform a prior task representation in order to perform a related new task zero-shot. The relationship by which the prior task embedding should be transformed to perform the new task is specified by example tuples of (input, output) task embeddings. This is precisely analogous to how a basic task is specified in terms of (input, output) embeddings of data points, except that the embeddings now represent tasks that are being transformed, rather than individual data points. \par
Thus from the perspective of our architecture, learning a meta-mapping between tasks is exactly analogous to learning a basic task. Anything that is embedded in $Z$ can be transformed using the same system. Because tasks are embedded in $Z$ for basic meta-learning, this allows for meta-mappings using exactly the same $\mathcal{M}$ and $\mathcal{H}$ that we use for basic tasks. Just as we would take a set of paired embeddings of data points for a basic task, and use them to compute a function embedding for that task, we can take a set of paired function embeddings representing a meta-mapping, and use them to create an embedding for that meta-mapping. We can then use this meta-mapping embedding to transform the embedding of another basic task. We can thus behave zero-shot on a novel task based on its relationship to a prior task. \par
For example, suppose we have an embedding $z_{game1} \in Z$ for the task of playing some game, and we want to switch to trying to lose this game. We can generate a meta-mapping embedding $z_{\text{meta}} \in Z$ from examples of embeddings generated by the system when it is trying to win and lose various other games: $z_{\text{meta}} = \mathcal{M}\left( \left\{\left((z_{game2},z_{game2,lose}\right), ... \right\}\right)$. We can generate a new task embedding $\hat{z}_{game1,lose} \in Z$:  

@@

\paragraph{Related work:} There has been a variety of past work on zero-shot learning, mostly based on natural-language descriptions of tasks. For example, \citet{Larochelle2008} considered the general problem of behaving in accordance with language instructions as simply asking a model to adapt its response when conditioned on different ``instruction'' inputs. Later work explored zero-shot classification based on only a natural language description of the target class \citep{Socher2013,Romera-Paredes2015,Xian2018}, or of a novel task in a language-conditioned RL system \citep{Hermann2017}. Some of this work has even exploited relationships between tasks as a learning signal \citep{Oh2017a}. Other work has considered how similarity between tasks can be useful for generating representations for a new task \citep{Pal2019}, but without transforming task representations to do so. (Furthermore, similarity cannot specify as rich of functions as an input-output mapping.) To our knowledge none of the prior work has proposed using meta-mapping approaches to adapt to new tasks by transforming task representations, nor has the prior work proposed a parsimonious homoiconic model which can perform these mappings. \par
